```{r setup}
#| message: false
#| include: false
library(tidyverse)
library(knitr)
library(cmdstanr)
library(bayesplot)
library(loo)
```

# Models {.unnumbered}

To do:

-   Add schemes of LTP & STP trajs with pars on it.
-   We need use random effect for control plots (plotoldnum).

## LTP

Long term process only model:

$$
\begin{align}
\mu_{rec} = \theta_{0,p} + (\theta_{\infty,s} - \theta_{0,p}) \times LTP \\
LTP = 1 - e^{-\lambda_p \times time}
\end{align}
$$

## STP

Long term and short term processes model based on disturbance index $dist$:

$$
\begin{align}
\mu_{rec} = \theta_{0,p} + (\theta_{\infty,s} - \theta_{0,p}) \times (LTP+STP) \\
LTP = 1 - e^{-\lambda_p \times time} \\
STP = \delta_p \times (\frac{time}{\tau_s} \times e^{1-\frac{time}{\tau_s}})^2 \\
\theta_{0,p} = \theta_{\infty,s} \times dist_p
\end{align}
$$

```{r stp_proj}
#| fig-cap: "STP proj with pars."
ltp <- function(time, dist, thetainf, lambda, tau, delta) {
  theta0 <- thetainf * dist
  ltp <- 1 - exp(-lambda * time)
  return(theta0 + (thetainf - theta0) * (ltp))
}
stp <- function(time, dist, thetainf, lambda, tau, delta) {
  theta0 <- thetainf * dist
  ltp <- 1 - exp(-lambda * time)
  stp <- delta * (time / tau * exp(1 - time / tau))^2
  return(theta0 + (thetainf - theta0) * (ltp + stp))
}
data.frame(time = seq(0.1, 50, length.out = 100)) %>%
  mutate(y_stp = stp(time, dist = 0.5, thetainf = 20, lambda = 0.05,
                     tau = 5, delta = 1)) %>%
  mutate(y_ltp = ltp(time, dist = 0.5, thetainf = 20, lambda = 0.05,
                     tau = 5, delta = 1)) %>%
  bind_rows(data.frame(time = 0, y_stp = 20, y_ltp = 20)) %>%
  ggplot(aes(time)) +
  geom_line(aes(y = y_ltp), col = "grey") +
  geom_line(aes(y = y_stp)) +
  theme_bw() +
  annotate("text", x = 50, y = 20, label = "theta[Inf]", parse = TRUE) +
  annotate("text", x = 20, y = 15, label = "lambda", parse = TRUE) +
  annotate("text", x = -1.5, y = 15, label = "dist", parse = TRUE) +
  annotate("text", x = 6, y = 10, label = "tau", parse = TRUE) +
  annotate("text", x = 6, y = 21, label = "delta", parse = TRUE) +
  theme(axis.title = element_blank())
```

Choices:

-   Using the log of y and a normal distribution centred on log(mu)
-   Cauchy hyperdist
-   Forced dist_p\<1
-   Bounds of theta
-   Free tau
-   Free delta

⚠️ Hump is mathematically opposed to the sign of dist_p-1.

## Old

### STP sum

`models/stp_sum.stan`

$$
\begin{align}
\mu_{rec} = \theta_{0,p} + (\theta_{\infty,s} - \theta_{0,p}) \times (LTP+STP) \\
LTP = 1 - e^{-\lambda_p \times time} \\
STP = \delta_p \times (\frac{time}{\tau_s} \times e^{1-\frac{time}{\tau_s}})^2 \\
\theta_{0,p} = \theta_{\infty,s} \times dist_p
\end{align}
$$

### STP lognormal

`models/stp.stan`

$$
\begin{align}
\mu_{rec} = \theta_{0,p} + (\theta_{\infty,s} - \theta_{0,p}) \times (LTP+STP) \\
LTP = 1 - e^{-\lambda_p \times time} \\
STP = \delta_p \times (\frac{time}{\tau_s} \times e^{1-\frac{time}{\tau_s}})^2 \\
\theta_{0,p} = \theta_{\infty,s} \times dist_p
\end{align}
$$

### STP correlated

`models/stp_cor.stan`

Imposed variance-covariance matrix on dist_p and delta_p.

$$
\begin{align}
\mu_{rec} = \theta_{0,p} + (\theta_{\infty,s} - \theta_{0,p}) \times (LTP+STP) \\
LTP = 1 - e^{-\lambda_p \times time} \\
STP = \delta_p \times (\frac{time}{\tau_s} \times e^{1-\frac{time}{\tau_s}})^2 \\
\theta_{0,p} = \theta_{\infty,s} \times dist_p
\end{align}
$$

## Parrallel

`dummy_par.stan`
