```{r set}
#| message: false
#| include: false
library(tidyverse)
library(knitr)
library(cmdstanr)
library(bayesplot)
library(loo)
library(patchwork)
list.files("r", full.names = TRUE, pattern = ".R") %>%
  lapply(source)
```

# Div - Genus q=1 {.unnumbered}

Genus diversity is measured using Hill's number with order $q=1$. Hill's numbers are parameterised by a diversity order $q$, which determines the sensitivity of the measures to the relative abundances of species. Hill's numbers include the three most commonly used measures of species diversity as special cases: species richness (q = 0), Shannon diversity (q = 1) and Simpson diversity (q = 2). The indices were derived from taxonomic censuses and cleaned taxonomies using the R package entropart [@marcon2015]. Finally, to account for differences in species identification, the indices have been corrected by a rarefaction curve, taking into account the value at 80% coverage.

```{r data}
#| message: false
#| warning: false
var <- "diversity_q1_gen"
lab <- expression("Genus diversity" ~ q == 1)
data <- load_all(var, lab)
```

## OGF

```{r ogf_data}
#| message: false
#| warning: false
#| fig-cap: "Old growth forests data."
ogf_data_fig(data)
```

```{r theta_ogf}
#| message: false
#| warning: false
#| fig-cap: "Theta posterior from old growth forests."
ogf_theta_fig(data)
```

## STP

```{r load_stp}
#| message: false
#| warning: false
data$fit_stp
```

```{r pvo_stp}
#| message: false
#| warning: false
#| fig-cap: "Predicted versus observed."
pvo_fig(data)
```

```{r preds_stp}
#| message: false
#| warning: false
#| fig-cap: "Predicted trajectories."
proj_fig(data)
```

```{r theta_stp}
#| message: false
#| warning: false
#| fig-cap: "Theta posterior from recovery trajectories against its posterior."
rec_theta_fig(data)
```
