```{r set}
#| message: false
#| include: false
library(tidyverse)
library(knitr)
```

# Group 2 {.unnumbered}

Group 2 blabla. We started from 3 questions:

-   **Q1**. How do forest attributes vary in recovery rates in the long-term?
-   **Q2**. How does recovery time vary with disturbance intensities?
-   **Q3**. How do abiotic and biotic factors modify recovery rates?

## Q1 - Recovery rates

> Q1. How do forest attributes vary in recovery rates in the long-term?

```{r lambda_a}
#| message: false
#| warning: false
#| fig-cap: "lambda across sites"
read_tsv("outputs/parameters.tsv") %>% 
  filter(level == "inter", 
         parameter != "lp__", parameter == "mu_lambda") %>% 
  ggplot(aes(x = attribute, col = type)) +
  geom_point(aes(y = median)) +
  geom_segment(aes(y = median-sd, yend = median+sd)) +
  geom_segment(aes(y = q5, yend = q95), size = .1) +
  coord_flip() +
  theme_bw() +
  ylab(expression(mu[~lambda])) + xlab("")
```

```{r t95_a}
#| message: false
#| warning: false
#| fig-cap: "t95 across sites"
read_tsv("outputs/parameters.tsv") %>% 
  filter(level == "inter", 
         parameter != "lp__", parameter == "mu_lambda") %>%
  mutate_at(c("median", "sd", "q5", "q95"), ~ log(20)/.) %>% 
  mutate_at(c("median", "sd", "q5", "q95"), ~ ifelse(. < 0, 0, .)) %>% 
  ggplot(aes(x = attribute, col = type)) +
  geom_point(aes(y = median)) +
  geom_segment(aes(y = median-sd, yend = median+sd)) +
  geom_segment(aes(y = q5, yend = q95), size = .1) +
  coord_flip() +
  theme_bw() +
  ylab(expression(mu[~t95])) + xlab("") +
  scale_y_log10()
```

## Q2 - Recovery time & intensity

> Q2. How does recovery time vary with disturbance intensities?

```{r lambda_dist_ba}
#| message: false
#| warning: false
#| fig-cap: "lambda across depending on dist BA"
dist_p <- read_tsv("outputs/parameters.tsv") %>% 
  filter(parameter == "dist", attribute == "ba") %>% 
  select(site, plot, median, sd, q5, q95) %>% 
  unique() %>% 
  rename_at(c("median", "sd", "q5", "q95"), ~ paste0(., "_dist_ba"))
read_tsv("outputs/parameters.tsv") %>% 
  filter(parameter == "lambda") %>% 
  left_join(dist_p) %>% 
  ggplot(aes(median_dist_ba, median)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point() +
  facet_wrap(~ attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  ggpubr::stat_cor(col = "red") +
  xlab(expression(dist[BA])) +
  ylab(expression(lambda))
```

```{r t95_dist_ba}
#| message: false
#| warning: false
#| fig-cap: "t95 across depending on dist BA"
dist_p <- read_tsv("outputs/parameters.tsv") %>% 
  filter(parameter == "dist", attribute == "ba") %>% 
  select(site, plot, median, sd, q5, q95) %>% 
  unique() %>% 
  rename_at(c("median", "sd", "q5", "q95"), ~ paste0(., "_dist_ba"))
read_tsv("outputs/parameters.tsv") %>% 
  filter(parameter == "t95") %>% 
  left_join(dist_p) %>% 
  ggplot(aes(median_dist_ba, median)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point() +
  facet_wrap(~ attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  ggpubr::stat_cor(col = "red") +
  xlab(expression(dist[BA])) +
  ylab(expression(t[95]))
```

## Q3 - Recovery rates & abiotic and biotic factors

> Q3. How do abiotic and biotic factors modify recovery rates?

abiotic factors = soil fertility + climate stress + landscape connectivity

biotic factors = biodiversity + Ecological strategy (acquisitive versus conservative) + structural diversity

lambda_p versus soil / climate / landscape variables (site in color)

lambda_p versus diversity_s/SLA_s/Gini_s of the controls/prelogging (site in color) 

T95_p versus soil / climate / landscape variables (site in color)

T95_p versus diversity_s/SLA_s/Gini_s of the controls/prelogging (site in color)
