```{r set}
#| message: false
#| include: false
library(tidyverse)
library(knitr)
```

# Group 2 {.unnumbered}

Group 2 blabla. We started from 3 questions:

-   **Q1**. How do forest attributes vary in recovery rates in the long-term?
-   **Q2**. How does recovery time vary with disturbance intensities?
-   **Q3**. How do abiotic and biotic factors modify recovery rates?

## Q1 - Recovery rates

> Q1. How do forest attributes vary in recovery rates in the long-term?

```{r lambda_a}
#| message: false
#| warning: false
read_tsv("outputs/parameters.tsv") %>%
  filter(
    level == "inter",
    parameter != "lp__", parameter == "mu_lambda"
  ) %>%
  ggplot(aes(x = attr_long, col = type)) +
  geom_point(aes(y = median)) +
  geom_segment(aes(y = median - sd, yend = median + sd)) +
  geom_segment(aes(y = q5, yend = q95), size = .1) +
  coord_flip() +
  theme_bw() +
  ylab(expression(mu[~lambda])) +
  xlab("") +
  ylim(0, NA)
```

```{r t90_a}
#| message: false
#| warning: false
read_tsv("outputs/parameters.tsv") %>%
  filter(
    level == "inter",
    parameter != "lp__", parameter == "mu_t90"
  ) %>%
  ggplot(aes(x = attr_long, col = type)) +
  geom_point(aes(y = median)) +
  geom_segment(aes(y = median - sd, yend = median + sd)) +
  geom_segment(aes(y = q5, yend = q95), size = .1) +
  coord_flip() +
  theme_bw() +
  ylab(expression(mu[~t90] ~ "[" ~ years ~ "]")) +
  xlab("") +
  scale_y_log10()
```

## Q2 - Recovery time & intensity

> Q2. How does recovery time vary with disturbance intensities?

```{r lambda_dist_ba}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
dist_p <- read_tsv("outputs/parameters.tsv") %>%
  filter(parameter == "phipct", attribute == "ba") %>%
  select(site, plot, median, sd, q5, q95) %>%
  unique() %>%
  rename_at(c("median", "sd", "q5", "q95"), ~ paste0(., "_dist_ba"))
read_tsv("outputs/parameters.tsv") %>%
  filter(parameter == "lambda") %>%
  left_join(dist_p) %>%
  ggplot(aes(median_dist_ba, median, col = site)) +
  geom_smooth(method = "lm", se = FALSE, col = "black", aes(group = NA)) +
  geom_point() +
  facet_wrap(~attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  xlab(expression(phi[BA] ~ "[ % lost ]")) +
  ylab(expression(lambda)) +
  scale_color_discrete("") +
  theme(legend.position = "bottom")
```

```{r t90_dist_ba}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
read_tsv("outputs/parameters.tsv") %>%
  filter(parameter == "t90") %>%
  left_join(dist_p) %>%
  ggplot(aes(median_dist_ba, median, col = site)) +
  geom_smooth(method = "lm", se = FALSE, col = "black", aes(group = NA)) +
  geom_point() +
  facet_wrap(~attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  xlab(expression(phi[BA] ~ "[ % lost ]")) +
  ylab(expression(t[90] ~ "[" ~ years ~ "]")) +
  scale_color_discrete("") +
  theme(legend.position = "bottom")
```

## Q3 - Recovery rates & abiotic and biotic factors

> Q3. How do abiotic and biotic factors modify recovery rates?

abiotic factors = soil fertility + climate stress + landscape connectivity

biotic factors = biodiversity + Ecological strategy (acquisitive versus conservative) + structural diversity

### Abiotic

```{r t90_cwd}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
read_tsv("outputs/parameters.tsv") %>%
  filter(parameter == "t90") %>%
  ggplot(aes(cwd, median)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(aes(col = site)) +
  facet_wrap(~attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  ylab(expression(t[90] ~ "[" ~ years ~ "]")) +
  scale_x_log10() +
  scale_color_discrete("") +
  theme(legend.position = "bottom") +
  xlab(expression(Climate ~ Water ~ Deficit ~ "[" ~ mm ~ year^{
    -1
  } ~ "]"))
```

```{r t90_tmax}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
read_tsv("outputs/parameters.tsv") %>%
  filter(parameter == "t90") %>%
  ggplot(aes(tmax, median)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(aes(col = site)) +
  facet_wrap(~attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  ylab(expression(t[90] ~ "[" ~ years ~ "]")) +
  scale_color_discrete("") +
  theme(legend.position = "bottom") +
  xlab(expression(Maximum ~ Temperature ~ "[" ~ "°" ~ "]"))
```

```{r t90_bdod}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
read_tsv("outputs/parameters.tsv") %>%
  filter(parameter == "t90") %>%
  ggplot(aes(bdod, median)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(aes(col = site)) +
  facet_wrap(~attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  ylab(expression(t[90] ~ "[" ~ years ~ "]")) +
  scale_color_discrete("") +
  theme(legend.position = "bottom") +
  xlab(expression(Bulk ~ Density ~ "[" ~ kg ~ dm^{
    -3
  } ~ "]"))
```

```{r t90_soc}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
read_tsv("outputs/parameters.tsv") %>%
  filter(parameter == "t90") %>%
  ggplot(aes(soc, median)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(aes(col = site)) +
  facet_wrap(~attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  ylab(expression(t[90] ~ "[" ~ years ~ "]")) +
  scale_color_discrete("") +
  theme(legend.position = "bottom") +
  xlab(expression(Soil ~ organic ~ Carbon ~ "[" ~ g ~ kg^{
    -1
  } ~ "]"))
```

```{r t90_deforested}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
read_tsv("outputs/parameters.tsv") %>%
  filter(parameter == "t90") %>%
  ggplot(aes(deforested, median)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(aes(col = site)) +
  facet_wrap(~attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  ylab(expression(t[90] ~ "[" ~ years ~ "]")) +
  scale_color_discrete("") +
  theme(legend.position = "bottom") +
  xlab(expression(Deforested ~ landscape ~ "[" ~ "%" ~ "]")) +
  scale_x_log10()
```

### Biotic

All site variables are the mean values for site $s$ across controls and pre-logging inventories.

```{r t90_gini}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
read_tsv("outputs/parameters.tsv") %>%
  filter(parameter == "t90") %>%
  ggplot(aes(gini_s, median)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(aes(col = site)) +
  facet_wrap(~attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  ylab(expression(t[90] ~ "[" ~ years ~ "]")) +
  xlab(expression(Gini[s])) +
  scale_color_discrete("") +
  theme(legend.position = "bottom")
```

```{r t90_sla}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
read_tsv("outputs/parameters.tsv") %>%
  filter(parameter == "t90") %>%
  ggplot(aes(SLA_cwm_ba_s, median)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(aes(col = site)) +
  facet_wrap(~attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  ylab(expression(t[90] ~ "[" ~ years ~ "]")) +
  xlab(expression(SLA ~ "[" ~ mm^2 ~ mg^{
    -1
  } ~ "]")) +
  scale_color_discrete("") +
  theme(legend.position = "bottom")
```

```{r t90_wd}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
read_tsv("outputs/parameters.tsv") %>%
  filter(parameter == "t90") %>%
  ggplot(aes(WD_cwm_ba_s, median)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(aes(col = site)) +
  facet_wrap(~attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  ylab(expression(t[90] ~ "[" ~ years ~ "]")) +
  xlab(expression(WD ~ "[" ~ g ~ cm^{
    -3
  } ~ "]")) +
  scale_color_discrete("") +
  theme(legend.position = "bottom")
```

```{r t90_gen}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
read_tsv("outputs/parameters.tsv") %>%
  filter(parameter == "t90") %>%
  ggplot(aes(diversity_q1_gen_s, median)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point(aes(col = site)) +
  facet_wrap(~attribute, scales = "free_y") +
  scale_y_log10() +
  theme_bw() +
  ylab(expression(t[90] ~ "[" ~ years ~ "]")) +
  xlab(expression("Genus diversity" ~ q == 1)) +
  scale_color_discrete("") +
  theme(legend.position = "bottom")
```
